---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Gaurav Girdhar"
date: "7/6/2018"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Section 1 - The Tidyverse Package

## Question 1

1. Can you name which package is associated with each task below?

Plotting - 

Data munging/wrangling -

Reshaping (speading and gathering) data -

Importing/exporting data -

## Answer 1

Plotting - ggplot

Data munging/wrangling - tidyr

Reshaping (spreading and gathering) data - tidyr

Importing & exporting Data - readr in tidyverse


## Question 2

Now can you name two functions that you’ve used from each package that you listed above for these tasks?

Plotting -

Data munging/wrangling -

Reshaping data -

Importing/exporting data -

## Answer 2

Plotting - geom_boxplot & geom_point

Data munging/wrangling - count & mutate

Reshaping data - gather & spread

Importing and Exporting Data - read_xxx & write_xxx

# Section 2 - R Basics

## Question 1

Fix this code with the fewest number of changes possible so it works:

My_data.name___is.too00ooLong! <- c( 1 , 2   , 3 )

## Answer 1

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
My_data.name___is.too00ooLong
```

## Question 2

Fix this code so it works:

my_string <- C('has', 'an', 'error', 'in', 'it)

## Answer 2

Changed 'C' to 'c' & included the missing closing '. 

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
my_string
```

## Question 3

Look at the code below and comment on what happened to the values in the vector.

my_vector <- c(1, 2, '3', '4', 5)

my_vector


## Answer 3

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

All the values are converted to Character by c()

# Section 3 - Data import/export

## Question 1

Download the rail_trail.txt file from Canvas (in the Midterm Exam section) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.

```{r}
library(tidyverse)
file_path <- "rail_trail.txt"
railtrails <- read_delim(file = file_path, delim = "|")
glimpse(railtrails)
```


## Question 2

Export the file into a comma-separated file and name it “rail_trail.csv”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse

## Answer 2

```{r}
write_csv(railtrails, path = 'rail_trail.csv')
file_path <- 'rail_trail.csv'
railtrails <- read_csv(file = file_path)
glimpse(railtrails)
```
Note : Filepath is set automatically by R Studio to the directory where .Rmd is.


# Section 4 - Visualization

## Question 1

Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.

## Answer 1

1) Dataset for each category is incomplete
2) Categories with gender would be more prominent
3) Response categories are not complete (Yes/No and expected the other categories too)

## Question 2

Reproduce this graphic using the diamonds data set.

## Answer 2

```{r}
library(tidyverse)
ggplot(data = diamonds, mapping = aes(factor(cut), carat, fill =  color)) + 
  geom_boxplot(position = "identity") +
  xlab ("CARAT OF DIAMOND") +
  ylab ("CUT OF DIAMOND") +
  coord_flip()
```


## Question 3

The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.


## Answer 3

Remove position parameter from boxplot

```{r}
library(tidyverse)
ggplot(data = diamonds, mapping = aes(factor(cut), carat, fill =  color)) + 
  geom_boxplot() +
  xlab ("CARAT OF DIAMOND") +
  ylab ("CUT OF DIAMOND") +
  coord_flip()
```


# Section 5 - Data Munging and Wrangling


## Question 1

Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.

```{r}
table2
```


## Answer 1

The data is not tidy. 

```{r}
table2 %>%
    spread(key = type, value = count)
```


## Question 2

Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.

## Answer 2

transmute (diamonds,
  price_per_carat = price/carat)
  
```{r}
transmute (diamonds,
  price_per_carat = price/carat)
```
  
  
## Question 3

### Question 3.1.1

For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.

### Answer 3.1.1
   
```{r}
diamonds %>%
   filter(carat < 1.5 & price > 10000) %>%
   count(cut) %>%
   mutate(prop = n / sum(n)) %>%
   print(n = 30)
```


### Question 3.1.2

Do the results make sense? Why?

### Answer 3.1.2

The reult does make sense as all the number in each catogory are increasing as per the cut.

### Question 3.1.3

Do we need to be wary of any of these numbers? Why?

### Answer 3.1.3

Very Good and Premium are very close.


# Section 4 - EDA

Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:

```{r}
glimpse(txhousing)
```


## Question 1

During what time period is this data from?

## Answer 1

The Data is between 2000 & 2015

## Question 2

How many cities are represented?

## Answer 2

```{r}
txhousing %>%
  count(city)
```

There are total 46 Cities

## Question 3

Which city, month and year had the highest number of sales?

## Answer 3

Texas <- txhousing %>%
  select(city, month, year, sales) %>%
  max.col(Texas) %>%
  print(Texas)


## Question 4

What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.

## Answer 4



## Question 5

What proportion of sales is missing for each city?

## Answer 5

```{r}
txhousing %>%
   filter(is.na(sales)) %>%
   count(city) %>%
   mutate(prop = n / sum(n)) %>%
   print(n = 30)
```

## Question 6

Looking at only the cities and months with greater than 500 sales:

```{r}
txhousing %>%
   filter(sales > 500) %>%
   select(city, month, sales)
```


Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.



Any cities that stand out that you’d want to investigate further?


Why might we want to filter out all cities and months with sales less than 500?