---
title: "COMPSCIX 415.2 Homework1"
author: "Gaurav Girdhar"
date: "June 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: https://github.com/ggirdhar28/compscix-415-2-assignments.git

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

Loading Data

```{r}
data("WorldCities")
```

QUESTION 1: How many observations and variables are there in this data set? What are some of the variable names? Type up your answer in your document using complete sentences.

ANSWER1: There are 23,018 Observations and 10 variable in this data set. Some of the Variable Names are as below:

a) Code
b) Name
c) Latitude
d) Longitude
e) Country
f) Country Region

```{r}
glimpse(WorldCities)
```

```{r}
WorldCities <- head(WorldCities, 200) # 200 rows
```
```{r}
country_col <- WorldCities$country
```
```{r}
unique(country_col)
```
QUESTION 2: There is a column called region in the data set. Can you extract this and show only the unique values?
```{r}
region_col <- WorldCities$region
```
```{r}
unique(region_col)
```
